Against the Planetary Interpretation of PTFO$\,$8-8695b

### 2020-03-28 paper/ms.tex
add figures and paper pdf
... and compressed the corner plot
add works for blend analysis

\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% ", Explaining its Decreasing Orbital Period"
% ": A Cautionary Tale in the Search for Tidal Orbital Decay"
% "An Explanation for WASP-4's Decreasing Orbital Period".
% "WASP-4's Decreasing Orbital Period: Solved."

@@

\begin{abstract}
  PTFO$\,$8-8695b could be the youngest, shortest-period 
  hot Jupiter known.  However it has not been shown to be a planet.
  TESS recently observed PTFO$\,$8-8695 for one month.
  The TESS light-curve shows that the dominant variability in this
  system is a sinusoidal modulation with period 11.96 hours, likely
  caused by stellar rotation.
  Also present is a complex signal, previously identified as the
  planet candidate, that repeats every 10.74 hours.
  The ``rotation'' and ``orbit'' have different frequencies,
  and show the expected beat every 4.48 days.
  There is a dip in the complex, shorter-period signal.
  However the orbital phase of the dip has changed over the past
  decade, in agreement with recent observations by Tanimoto et al.
  Planets do not change their orbital phase.
  PTFO$\,$8-8695 therefore seems broadly consistent with the
  ``transient dipping'' phenomenology observed in many young M dwarfs,
  and seems unlikely to be a planet.
\end{abstract}

@@

% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.7\textwidth]{f1.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
% 		The vertical axis shows the observed transit times minus the
% 		calculated times, assuming a constant orbital period.  More opaque
% 		points correspond to more precise data.  The $\pm1\sigma$
% 		uncertainties of the quadratic ephemeris are shown in blue.  The
% 		yellow star represents the weighted average of 18 data points
% 		obtained with TESS. The TESS data were averaged here for display
% 		purposes only; our analysis used the 18 individual transit times.
% 		\label{fig:times}
% 	}
% \end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:lc_map}
	}
\end{figure*}

\begin{figure*}[t]
	% \begin{center}
	% 	\leavevmode
	% 	\includegraphics[width=0.95\textwidth]{f1.png}
	% \end{center}
  \gridline{
    \fig{f2a.png}{0.8\textwidth}{}
    \fig{f2b.png}{0.2\textwidth}{}
  }
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:splitsignal}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f3.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:phasefold}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:corner}
	}
\end{figure*}

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  % \texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

>>>>

\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}

@@

\begin{abstract}
  PTFO$\,$8-8695b could be the youngest, shortest-period 
  hot Jupiter known.  However it has not been shown to be a planet.
  TESS recently observed PTFO$\,$8-8695 for one month.
  The TESS light-curve shows that the dominant variability in this
  system is a sinusoidal modulation with a ``long'' period $P_{\rm
  \ell}$
  of 11.96 hours, likely caused by stellar rotation.
  Also present is a complex signal, previously identified as the
  planet candidate, that repeats with a ``short'' period $P_{\rm s}$ of
  10.74 hours.
  The ``long'' and ``short'' signals show the expected beat every 4.48
  days.
  There is a dip in the complex, short-period signal.
  However ground-based photometry from the past decade shows that the
  orbital phase of the dip seems to have instantaneously jumped, at
  least twice.
  The TESS epoch of the dip is consistent with recent observations by
  Tanimoto et al., and differs from the discovery epoch by 5.14 hours.
  Planets do not ``jump'' in orbital phase.
  PTFO$\,$8-8695 therefore seems broadly consistent with the
  ``transient dipping'' phenomenology observed in many young M dwarfs,
  and seems unlikely to be a planet.
\end{abstract}

@@

\subsection{Blend considerations}
\label{subsec:blend}

The TESS pixels are $\approx21$'' per side, and so we need to consider
whether light from neighboring stars could affect the photometry.  The
scene is shown in Figure~\ref{fig:scene}.  
The pixels used to
measure the background level are indicated with an `\texttt{X}' hatch,
and the pixels used for the final light-curve are shown with the
`\texttt{/}' hatch.

The target star, PTFO
8-8605 (TIC 264461976), has a $T$-band magnitude of 14.0, and its position is shown with a
star.  
The other (unlabeled) star inside the target aperture, TIC 264461979, has $T=16.8$ and so cannot
contribute a signal with relative amplitude 10\%.
The only neighbor that is sufficiently close and bright that
its light might contaminate the target star is TIC 264461980, with
$T=14.8$, which we denote ``Star A''.  Star A is 23.6'' NW of our
target, and based on the magnitude difference could contribute up to
48\% the flux of our target star, PTFO 8-8695.

Because PTFO 8-8695 has previously been identified to have periodicity
consistent with our measurement of $P_{\rm s}$, our main concern
regarding blending is the degree to which we can be certain that the
long-period signal at $P_{\rm \ell}$ also originates from PTFO 8-8695.
We took two approaches towards determining the source of the long-period signal.

First, we examined the CDIPS full frame image light-curves of the
target, which are available on MAST \citep{bouma_cluster_2019}.
The maximal peak-to-peak beat amplitude is consistently $\approx$10\%
across apertures of radii 1, 1.5, and 2.25 pixels.
If Star A were the source of the long-period variability, we would expect the
peak variability amplitude to be smallest in the 1 pixel aperture, based on the
separation of the sources (Figure~\ref{fig:scene}, bottom).
From this test alone, it seems unlikely that Star A is the source of
the long-period signal.

Second, we examined the light-curve of each pixel in the scene
individually.  We opted to use the
interactive tools implemented in
\texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
source of the long-period variability, we would expect the pixels
nearest to Star A to show a sinusoidal signal with
amplitude exceeding $10\%$.  We find no evidence for
this being the case.  The pixel directly below Star A does not
clearly show the sinusoidal variability, and the peak-to-peak 
variability in that pixel is $\lesssim 8\%$.  In contrast, the
south-easternmost pixel within PTFO 8-8695's aperture (the pixel 
furthest from Star A that was used in the optimal aperture) shows the $P_{\rm \ell}$ sinusoidal
variability signal at $\approx 10\%$ amplitude.

As there is no evidence in favor of a blend scenario, we
conclude that both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from PTFO 8-8695.

@@

%\begin{figure*}[t]
%	\begin{center}
%		\leavevmode
%		\includegraphics[width=0.95\textwidth]{f2.png}
%	\end{center}
%	\vspace{-0.7cm}
%	\caption{ {\bf foo.}
%    bar
%		\label{fig:lc_map}
%	}
%\end{figure*}

\begin{figure*}[t]
  \gridline{
    \fig{f3a.png}{0.8\textwidth}{}
    \fig{f3b.png}{0.2\textwidth}{}
  }
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:splitsignal}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f4.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:phasefold}
	}
\end{figure*}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene used for blend analysis.}
		{\it Top:} Mean TESS image of PTFO 8-8695 over Sector~6, with a log-stretch.
		The position of PTFO 8-8695 is shown with a yellow star.
		Neighbors with $T<17$ are shown with red crosses.
		The apertures used to measure the background and target star flux
		are shown with \texttt{X} and \texttt{/} hatches, respectively.
		{\it Bottom:} DDS2 $R$-band image of the same field, with a linear stretch. The circles show
		apertures of radii 1, 1.5, and 2.25 pixels used in part of our blend
		analysis.
		The pixel level TESS data show that ``Star A''  does not contribute variability at either of the two observed periods (see Section~\ref{subsec:blend}).
		\label{fig:scene}
	}
\end{figure}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{o_minus_c.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
		bar
		\label{fig:o_minus_c}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f5_comp.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:corner}
	}
\end{figure*}

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  % \texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}
